---
title: "stm"
author: "Ryan Yeung"
date: '2024-12-11'
format:
  html:
    toc: true
    toc-location: left
    code-fold: true
    code-tools: true
    code-block-bg: true
    code-link: true
    theme: flatly
    df-print: paged
    embed-resources: true
    self-contained: true
---

## Structural Topic Models (STM)

```{r setup}
if (!require("pacman")) {
  install.packages("pacman")
}

pacman::p_load(
  ### uncomment the following line if data are in an excel sheet
  # readxl,
  tidyverse, gtsummary, janitor,
  stm,
  ### for textProcessor()
  tm, SnowballC,
  ### for K = 0 in stm()
  Rtsne, rsvd, geometry,
  ### for bonus preprocessing
  udpipe, quanteda, quanteda.textstats,
  kableExtra, pander
)
```

## Data Prep

::: {.panel-tabset}

### Import

```{r import}
df_raw <-
  read_csv("../input/dreaddit-train.csv")
  ### run the following line instead if data are in an excel sheet
  # readxl::read_excel("file_name_here.xlsx")

df_raw %>%
  head(10)
```

### Preprocessing

```{r preprocessing}
processed <- textProcessor(df_raw$text, metadata = df_raw)
out <- prepDocuments(processed$documents, processed$vocab, processed$meta)

summary(out)
```

:::

## Run Basic Model

::: {.panel-tabset}

### Set Up STM

```{r stm setup}
docs <- out$documents
vocab <- out$vocab
meta <- out$meta

cat("Number of Documents: ", length(docs))
summary(vocab)
summary(meta)
```

### Basic STM

```{r basic stm}
topic_model <-
  stm(
    documents = out$documents,
    vocab = out$vocab,
    K = 0,
    prevalence = ~subreddit + social_karma,
    max.em.its = 75,
    data = out$meta,
    init.type = "Spectral",
    seed = "111224"
    )

topic_model
```

```{r save basic topic model}
saveRDS(topic_model, "../output/stm_basic.rds")
```

### SearchK

```{r find k}
findingk <- 
  searchK(
    documents = out$documents,
    vocab = out$vocab,
    K = c(20:30),
    prevalence = ~subreddit + social_karma,
    data = meta,
    verbose = FALSE,
    )
```

```{r findingk diagnostics}
plot(findingk)

findingk$results %>%
  kable(
    table.attr = "style = \"color: black;\"", 
    caption = "summary statistics for simulated topic models"
    ) %>% 
  kableExtra::kable_styling(c("striped", "hover", "condensed")) %>%
  kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r findingk diagnostics ggplot}
### plot diagnostic values for all possible Ks
findingk$results %>% 
  mutate(across(.cols = everything(), as.numeric)) %>% 
  select(-bound, -em.its) %>% 
  pivot_longer(., cols = -K, names_to = "metric", values_to = "value") %>% 
  ggplot(aes(K, value, color = metric)) + 
  geom_line(linewidth = 0.5, alpha = 0.7, show.legend = FALSE) +
  geom_point(show.legend = FALSE) + 
  facet_wrap(~metric, scales = "free") +
  ggrepel::geom_text_repel(
    aes(label = K),
    show.legend = FALSE,
    ) +
  labs(
    x = "K (Number of Topics)",
    y = NULL,
    ) + 
  theme_classic()
```

:::

## Inspect Basic Model

```{r extract number of topics}
n_topics <- 
  topic_model$settings$dim$K %>% 
  as.numeric()

cat("Number of Topics: ", n_topics)
```

::: {.panel-tabset}

### Topic Prevalence

```{r topic prevalence}
df_topic_prev <-
  ### calculate mean thetas (proportion of document associated with a topic) across all documents
  colMeans(topic_model$theta[, 1:n_topics]) %>% 
  data.frame() %>% 
  rename("prevalence" = ".") %>% 
  mutate(
    topic = paste0("Topic ", row_number()),
    topic_num = row_number(),
    prevalence = round(prevalence*100, 2),
    ) %>%
  relocate(topic)

df_topic_prev

### identify top 10 most prevalent topics
top_n_prevalent_topics <- 
  df_topic_prev %>% 
  slice_max(prevalence, n = 20) %>% 
  select(topic_num) %>% 
  as.list() %>% 
  unlist(use.names = FALSE)

top_n_prevalent_topics
```

### Plot Topic Prevalence 

```{r plot topic prevalence}
plot(
  topic_model, width = 30, text.cex = 0.7,
  topics = top_n_prevalent_topics,
  labeltype = "frex", n = 5,
  main = "Topics by Prevalence and Top FREX Words"
  )
```

### Label Topics

```{r label topics}
labelTopics(topic_model, top_n_prevalent_topics)
```

### Representative Documents

```{r representative documents}
df_rep_doc <-
  data.frame()

for (i in top_n_prevalent_topics) {
  df <- 
    findThoughts(
      topic_model, 
      texts = meta$text, 
      topics = i,
      n = 5,
      )$docs[[1]] %>%
    as.data.frame() %>%
    mutate(topic_num = i) 
  
  df_rep_doc <-
    bind_rows(df_rep_doc, df)
}

df_rep_doc <-
  df_rep_doc %>% 
  rename("text" = ".")

df_rep_doc
```

:::

## Inference

::: {.panel-tabset}

### Estimate Effect

```{r estimate effect}
topic_model_est <-
  estimateEffect(
    formula = top_n_prevalent_topics ~ subreddit + social_karma,
    stmobj = topic_model,
    metadata = out$meta,
    uncertainty = "None"
    )
```

### Extract Effects Table

```{r extract estimate effects as table}
topic_model_est_effects <- summary(topic_model_est)

topic_model_est_effects_summary <- 
  ### bind summary outputs together
  do.call(rbind, topic_model_est_effects[[3]]) %>% 
  as.data.frame()

df_est_effects <-
  topic_model_est_effects_summary %>% 
  rownames_to_column(var = "predictor") %>% 
  filter(!str_detect(predictor, "Intercept")) %>% 
  janitor::clean_names()

df_est_effects$topic_num <-
  rep(top_n_prevalent_topics, each = 10)

df_est_effects %>% 
  select(-t_value) %>% 
  filter(pr_t < .05)
```

### Plot Effects

```{r}
### topics significantly related to 
for (i in c(34, 12, 1)) {
  plot(
    topic_model_est, covariate = "social_karma",
    topics = i, main = paste0("Topic ", i),
    model = topic_model, method = "continuous",
    xlab = "Karma", linecol = "blue",
    printlegend = FALSE
    )
}
```

```{r}
### 58 seems related to symptoms or treatment ("symptom", "nightmar", "depress")
### accordingly, related to both PTSD and anxiety subreddits
plot(
  topic_model_est, covariate = "subreddit",
  topics = c(51),
  model = topic_model, method = "pointestimate",
  xlab = "Prevalence", linecol = "blue",
  printlegend = FALSE
  )
```

```{r}
### 25 seems related to components of traumatic memory ("abus", "rememb", "memori")
### accordingly, related to trauma/PTSD subreddits as opposed to anxiety
plot(
  topic_model_est, covariate = "subreddit",
  topics = c(25),
  model = topic_model, method = "pointestimate",
  xlab = "Prevalence", linecol = "blue",
  printlegend = FALSE
  )
```

:::

## Bonus: Better Preprocessing

::: {.panel-tabset}

### UDPipe

```{r udpipe model}
udmodel <- udpipe_load_model(file = "../input/english-gum-ud-2.5-191206.udpipe")
udmodel
```

### Parse Text

```{r parse text}
df_parsed <-
  udpipe_annotate(
    object = udmodel,
    x = df_raw$text,
    doc_id = df_raw$id,
    ) %>%
  as.data.frame()

df_parsed %>% 
  head(50)
```

### Clean Parsed Text

```{r clean parsed text}
df_parsed <-
  df_parsed %>% 
  ### return doc_id to numeric (parsing converts to character)
  mutate(doc_id = as.numeric(doc_id)) %>% 
  ### remove punctuation, numbers, symbols
  filter(!(upos == "PUNCT")) %>% 
  filter(!(upos == "NUM")) %>% 
  filter(!(upos == "SYM"))

df_parsed %>% 
  select(upos) %>% 
  tbl_summary()

selected_vars <-
  c("subreddit", "social_karma") 

### join text metadata from df_pivot (`selected_vars`) back onto parsed df
df_parsed <- 
  df_raw %>% 
  ### select relevant metadata variables from text data CSV
  select(id, any_of(selected_vars)) %>% 
  mutate(doc_id = as.numeric(id)) %>% 
  right_join(df_parsed, by = join_by("doc_id"))

df_parsed %>% 
  slice_sample(n = 50)
```

### Concatenate Lemmas

```{r concat parsed lemmas}
df_parsed_concat <- 
  df_parsed %>% 
  group_by(doc_id) %>% 
  summarize(text_lemmas = paste(lemma, collapse = " ")) %>% 
  ungroup() %>% 
  rename("id" = "doc_id") %>% 
  left_join(x = df_raw, y = ., by = "id")

df_parsed_concat %>% 
  select(id, text_lemmas, text) %>% 
  head(20)
```

### Create Corpus & DFM

```{r create corpus}
df_corpus <-
  df_parsed_concat %>% 
  # as.data.frame() %>% 
  corpus(
    docid_field = "doc_id",
    text_field = "text_lemmas",
  )

df_corpus %>% 
  docvars() %>%
  head(20)
```

```{r create dfm}
df_dfm <-
  df_corpus %>% 
  tokens() %>% 
  dfm()

df_dfm
```

### Vocab Pruning

```{r vocab pruning}
df_dfm_freq <- textstat_frequency(df_dfm) # check high freq tokens for cutoffs

### frequency head and tail
df_dfm_freq %>%
  arrange(desc(frequency)) %>% 
  slice(1:50, (n()-49):n())

### docfreq head and tail
df_dfm_freq %>%
  arrange(desc(docfreq)) %>% 
  slice(1:50, (n()-49):n())

### docfreq distribution
df_dfm_freq %>% 
  # filter(docfreq > 100) %>%
  ggplot(aes(x = docfreq)) +
  # geom_histogram(binwidth = 5)
  geom_density() +
  scale_x_continuous(limits = c(0, 600, 100), breaks = seq(0, 600, 100))

tabyl(df_dfm_freq$docfreq) %>% 
  mutate(percent = round(percent*100, 2)) %>% 
  mutate(cumul_percent = cumsum(percent)) %>% 
  kableExtra::kable(table.attr = "style = \"color: black;\"", caption = "Document Frequency") %>% 
  kableExtra::kable_styling(c("striped", "hover", "condensed")) %>%
  kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r vocab pruning thresholds freq}
### 3SD method seems somewhat conservative
maxfreq_3SD <- 
  round(mean(df_dfm_freq$frequency) + (3*sd(df_dfm_freq$frequency))) 

minfreq_3SD <- 
  round(mean(df_dfm_freq$frequency) - (3*sd(df_dfm_freq$frequency)))

### some recommend min 1% doc frequency
min1pctdocfreq <- round(ndoc(df_dfm)/100) 
### max 99% doc frequency is also recommended
max99pctdocfreq <- round(99*ndoc(df_dfm)/100) 

### SD and percent freq thresholds in table
minmaxfreq_df <- data.frame(minfreq_3SD, maxfreq_3SD, min1pctdocfreq, max99pctdocfreq)

colnames(minmaxfreq_df) <- c("mean freq - 3 SDs", "mean freq + 3 SDs", "1% doc freq", "99% doc freq")

pander::pandoc.table(
  minmaxfreq_df, style = "rmarkdown",
  caption = "mean ± 3SD freq versus 1% and 99% docfreq thresholds",
  )
```

```{r trim dfm}
### trim dfm based on minimum 0.5% docfreq (>=3) and maximum +3SD freq (>=811)
df_dfm_trimmed <- dfm_trim(df_dfm, min_docfreq = 3, max_termfreq = 909, termfreq_type = "count")
df_dfm_shape <- c(ndoc(df_dfm), nfeat(df_dfm))
df_dfm_trimmed_shape <- c(ndoc(df_dfm_trimmed), nfeat(df_dfm_trimmed))
df_dfm_prepost_trim_shape <- data.frame(rbind(df_dfm_shape, df_dfm_trimmed_shape))

colnames(df_dfm_prepost_trim_shape) <- c("ndoc", "nfeat")

pander::pander(
  df_dfm_prepost_trim_shape, style = "rmarkdown",
  caption = "DFM shape before and after vocab pruning (min doc freq >= 3, max term freq <= 909)"
  )

### recheck high freq tokens for cutoffs
df_dfm_trimmed_freq <- textstat_frequency(df_dfm_trimmed) 

### frequency head and tail
df_dfm_trimmed_freq %>%
  arrange(desc(frequency)) %>%
  slice(1:50, (n()-49):n())

### docfreq head and tail
df_dfm_trimmed_freq %>%
  arrange(desc(docfreq)) %>% 
  slice(1:50, (n()-49):n())
```

### STM Setup

```{r stm setup with trimmed}
out <- convert(df_dfm_trimmed, to = "stm")

documents <- out$documents
vocab <- out$vocab
meta <- out$meta

cat("Number of Documents: ", length(docs))
summary(vocab)
summary(meta)
```

:::

## Cleanup

::: {.panel-tabset}

### Save Model

```{r save rdata, warning = F}
rdata_file_name <- 
  paste0(
    "../stm_model/",
    as.character(format(Sys.time(), '%F')), ### add "_%I-%M%p"for time
    "_stm_basic.RData", sep = ""
    )

rdata_file_name
save.image(file = rdata_file_name, safe = TRUE)
```

### Cite Packages

```{r cite packages}
report::cite_packages()
```

:::